### 2018-11-19 Sections/quantum_algorithms.tex
tiny
Finished QFT box

If we consider the case of solving a satisfiability problem, to find a satisfying input such that the function evaluates to \mbox{$f(x)=1$}, then it would na{\"i}svely appear that the required quantum operation needs to implement the transformation,
\begin{align}
	\hat{U}_f\ket{x}=\ket{f(x)}.
\end{align}
However, it is easy to see that in general such a transformation is non-unitary, and therefore cannot be implemented quantum mechanically. To overcome this obstacle and enable unitary implementation, we introduce additional ancillary bits\index{Ancillary bits} to $f(x)$ and first write it as a reversible classical circuit\index{Reversible classical circuits}, which then always lends itself to a unitary quantum construction. For example, if we instead introduce some extra qubits and implement the transformation,
\begin{align}
\hat{U}_f\ket{x}\ket{0} 	= \ket{x}\ket{f(x)},
\end{align}
then it can easily be seen that this preserves orthonormality and may be implemented unitarily for any function $f(x)$, even if $f(x)$ is not an invertible function.

To provide the simplest possible example, consider the classical XOR gate, given by,
\begin{align}
	y_1 = x_1\oplus x_2.
\end{align}
This operation is obviously not reversible (and hence not unitary), since for a give output $y_1$, the inputs $x_1$ and $x_2$ are not unique. However, if we trivially modify the transformation to be,
\begin{align}
	y_1 &= x_1\oplus x_2,\nonumber\\
	y_2 &= x_2,
\end{align}
then a quick back-of-the-envelope calculation verifies that the transformation (simply a CNOT gate\index{CNOT gate} now) is now reversible, unitary, and output $y_1$ encodes the desired function evaluation.

Having re-expressed our desired oracle function as a reversible classical circuit, we now simply make direct substitutions of all the reversible classical gates with their logically equivalent quantum counterparts, yielding a quantum implementation of the function that is unitary and preserves quantum coherence. The progression is shown in Fig.~\ref{fig:oracle_circ}.

@@

@@

@@

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function QFT($\ket{x}$):
\begin{enumerate}
\item Apply the recursively-defined circuit below to the $n$-qubit input state $\ket{x}$, where,
\begin{align}
\hat{R}_\theta=\left(\begin{matrix}
1 & 0 \\
0 & e^{i\theta}	
\end{matrix}\right),
\end{align}
is a generalised phase-flip gate.
\item The amplitudes in the $n$-qubit output state $\ket{y}$ are given by the quantum Fourier transform of the input amplitudes, implementing the transformation,
\begin{align}
	\hat{\mathrm{QFT}} \ket{j} \to \frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} \omega^{jk}\ket{k},
\end{align}
on the logical basis states, or in matrix form,
\begin{align}
\frac{1}{\sqrt{N}} \left(\begin{matrix}
  1 & 1 & 1 & \dots & 1\\
  1 & \omega & \omega^2 & \dots & \omega^{N-1} \\
  1 & \omega^2 & \omega^4 & \dots & \omega^{2(N-1)} \\
  \vdots & \vdots & \vdots & \ddots & \vdots \\
  1 & \omega^{N-1} & \omega^{2(N-1)} &\dots &\omega^{(N-1)(N-1)}
\end{matrix}\right),
\end{align}
where there are,
\begin{align}
N=2^n,
\end{align}
logical basis states.
\item The quantum circuit requires $O(n^2)$ gates.
\item $\Box$
\end{enumerate}
\includegraphics[width=\textwidth]{QFT}
}
\end{mdframed}
\captionspacealg \caption{The quantum Fourier transform (QFT) algorithm. The circuit is defined recursively, defining the $n$-qubit QFT in terms of an \mbox{$n-1$}-qubit QFT.} \label{alg:QFT}\index{Quantum Fourier transform}
\end{table}

>>>>

If we consider the case of solving a satisfiability problem, to find a satisfying input such that the function evaluates to \mbox{$f(x)=1$}, then it would na{\"i}vely appear that the required quantum operation needs to implement the transformation,
\begin{align}\label{eq:wrong_oracle}
	\hat{U}_f\ket{x}=\ket{f(x)}.
\end{align}
However, it's easy to see that in general such a transformation is non-unitary, and therefore cannot be implemented quantum mechanically.

To overcome this obstacle and enable unitary implementation, we introduce additional ancillary bits\index{Ancillary bits} to $f(x)$, and first write it as a reversible classical circuit\index{Reversible classical circuits}, which then always lends itself to a unitary quantum construction. Specifically, if instead of attempting to construct Eq.~(\ref{eq:wrong_oracle}) we  introduce some extra qubits and attempt to implement the transformation,
\begin{align}
\hat{U}_f\ket{x}\ket{0} 	= \ket{x}\ket{f(x)},
\end{align}
then it can easily be seen that this preserves orthonormality and may be implemented unitarily for any function $f(x)$, even if $f(x)$ is not an invertible function.

To provide the simplest possible example, consider the classical XOR gate, given by,
\begin{align}
	y_1 = x_1\oplus x_2.
\end{align}
This operation is obviously not reversible (and hence not unitary), since for a given output $y_1$, the inputs $x_1$ and $x_2$ are not unique. However, if we trivially modify the transformation to be,
\begin{align}
	y_1 &= x_1\oplus x_2,\nonumber\\
	y_2 &= x_2,
\end{align}
then a quick back-of-the-envelope calculation verifies that the transformation (simply a CNOT gate\index{CNOT gate}) is now reversible, unitary, and output $y_1$ encodes the desired function evaluation.

Having re-expressed our desired oracle function as a reversible classical circuit, we now simply make direct substitutions of all the reversible classical gates with their logically equivalent quantum counterparts, yielding a quantum implementation of the function that is unitary and preserves quantum coherence. The progression for this example is shown in Fig.~\ref{fig:oracle_circ}. The general procedure for oracle construction is shown in Alg.~\ref{alg:oracle_constr}

@@

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function Oracle(f):
\begin{enumerate}
	\item Let $f$ be an arbitrary, efficiently-computable function in \textbf{BPP}.
	\item Express $f$ as a classical logic circuit.
	\item Make ancillary bits available.
	\item Rewrite the classical circuit as a reversible classical circuit, exploiting the introduced ancillary qubits as necessary.
	\item Make direct substitutions of all reversible classical gates with their quantum counterparts that implement the equivalent logical operations in qubit space.
	\item Return $\hat{U}_f$.
    \item $\Box$
\end{enumerate}
}
\end{mdframed}
\captionspacealg \caption{Outline of the general procedure for constructing quantum oracles from classical logic descriptions.} \label{alg:oracle_const}
\end{table}

@@

Note that while the circuit only requires $O(n^2)$ gates, it implements a Fourier transform on $2^n$ amplitudes, giving this subroutine an exponential quantum improvement over classical DFTs.

@@

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function QFT($\ket{x}$):
\begin{enumerate}
\item Apply the recursively-defined circuit below to the $n$-qubit input state $\ket{x}$, where,
\begin{align}
\hat{R}_\theta=\left(\begin{matrix}
1 & 0 \\
0 & e^{i\theta}	
\end{matrix}\right),
\end{align}
is a generalised phase-flip gate.
\item The amplitudes in the $n$-qubit output state $\ket{y}$ are given by the quantum Fourier transform of the input amplitudes, implementing the transformation,
\begin{align}
	\hat{\mathrm{QFT}} \ket{j} \to \frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} \omega^{jk}\ket{k},
\end{align}
on the logical basis states, or in matrix form,
\begin{align}
\frac{1}{\sqrt{N}} \left(\begin{matrix}
  1 & 1 & 1 & \dots & 1\\
  1 & \omega & \omega^2 & \dots & \omega^{N-1} \\
  1 & \omega^2 & \omega^4 & \dots & \omega^{2(N-1)} \\
  \vdots & \vdots & \vdots & \ddots & \vdots \\
  1 & \omega^{N-1} & \omega^{2(N-1)} &\dots &\omega^{(N-1)(N-1)}
\end{matrix}\right),
\end{align}
where there are,
\begin{align}
N=2^n,
\end{align}
logical basis states.
\item The quantum circuit requires $O(n^2)$ gates.
\item Return $\ket{y}$.
\item $\Box$
\end{enumerate}
\includegraphics[width=\textwidth]{QFT}
}
\end{mdframed}
\captionspacealg \caption{The quantum Fourier transform (QFT) algorithm. The circuit is defined recursively, defining the $n$-qubit QFT in terms of an \mbox{$n-1$}-qubit QFT.} \label{alg:QFT}\index{Quantum Fourier transform}
\end{table}
